---
title: "BIOS 735 Project Report"
author: "Group 2"
date: "4/6/2022"
output: html_document
---

Deliverable details:

Summarize results in Rmarkdown document, submitted separately from the R package. Aim for Introduction to the dataset and the question, a Methods section with any important pre-processing etc., three sections of Results (which may also contain some method details), and Discussion about interpretation or possible future directions. This will be due 5 pm, May 8th.

- due date is actually (Saturday April 30th by 11:59 PM)

```{r, echo=FALSE}
# set all code chunks to not render in output (only output & text)
knitr::opts_chunk$set(echo = FALSE)
```

```{r, warning=FALSE, message=FALSE}
# load relevant packages (including the package we develop)
#library(HeartDiseaseOptim) # our package
library(tidyverse)
library(mosaic)
library(caret)

# load heart disease dataset
heart.data <- read.csv("~/bios735_group2/data/heart_2020_cleaned.csv") # andrew
```


### Introduction

- the dataset and the question

* problem being addressed (estimation of heart disease prevalence)
* description of the dataset
        * outcome of interest
        * additional covariates
* description of project aims

### Methods

- Methods section with any important pre-processing etc.

### Results

- three sections (module 2 statistical model / comparison to module 3 method / implementation in R package, module 1)

### Discussion

- interpretation or possible future directions

* summary of main points, conclusions
* limitations and concerns
* future problems to address