---
title: "BIOS 735 Primary Project Code"
author: "Group 2"
date: "4/4/2022"
output: html_document
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE, warning=FALSE}
# load relevant packages
library(tidyverse)
library(optimx)
library(caret)
library(fastDummies)
library(glmnet)
```


### Expression for Log Likelihood

The logistic regression model of the heart disease and other variables:
$$ 
Logit(\boldsymbol{\beta}, \boldsymbol{X}) = log(\frac{\boldsymbol{p}}{1-\boldsymbol{p}}) = {\beta}_0 + \beta_{1}X_1 + \beta_{2}X_2 + ... + \beta_{18}X_{18}
$$
The $p$ denotes the probability of getting heart disease, then:
$$ 
 \boldsymbol{p} = \frac{e^{\boldsymbol{X}\boldsymbol{\beta}}}{1+ e^{\boldsymbol{X}\boldsymbol{\beta}}}
$$

The likelihood could be written as:

$$ 
L(\boldsymbol{\theta}) = \prod_{i=1}^n (p_i)^{I[y_i=1]}(1-p_i)^{I[y_i=0]}
$$
The log likelihood could be further written as:

$$ 
l(\boldsymbol{\theta}) = \sum_{i=1}^n log(p_i) y_i + log(1-p_i) (1-y_i)
$$

### Expression for logistic regression model with elastic net 

```{r}
# set a randomization seed
set.seed(735)

# load heard disease data (add your own filepath to access the data)
heart.data <- read.csv('C:/Users/faye__000/Documents/Graduate schools/UNC/BIOS735/heart_2020_cleaned.csv')
head(heart.data)
str(heart.data)

## data pre-processing
# set factors as factors
heart.data$HeartDisease <- as.factor(heart.data$HeartDisease)
heart.data$Smoking <- as.factor(heart.data$Smoking)
heart.data$AlcoholDrinking <- as.factor(heart.data$AlcoholDrinking)
heart.data$Stroke <- as.factor(heart.data$Stroke)
heart.data$DiffWalking <- as.factor(heart.data$DiffWalking)
heart.data$Sex <- as.factor(heart.data$Sex)
heart.data$AgeCategory <- as.factor(heart.data$AgeCategory)
heart.data$Race <- as.factor(heart.data$Race)
heart.data$Diabetic <- as.factor(heart.data$Diabetic)
heart.data$PhysicalActivity <- as.factor(heart.data$PhysicalActivity)
heart.data$GenHealth <- as.factor(heart.data$GenHealth)
heart.data$Asthma <- as.factor(heart.data$Asthma)
heart.data$KidneyDisease <- as.factor(heart.data$KidneyDisease)
heart.data$SkinCancer <- as.factor(heart.data$SkinCancer)

# set numerics as numeric
heart.data$BMI <- as.numeric(heart.data$BMI)
heart.data$PhysicalHealth <- as.numeric(heart.data$PhysicalHealth)
heart.data$MentalHealth <- as.numeric(heart.data$MentalHealth)
heart.data$SleepTime <- as.numeric(heart.data$SleepTime)

# create training data index
trainIndex <- createDataPartition(heart.data$HeartDisease, p = .8, list = FALSE, times = 1)
# train set
heart.train <- heart.data[trainIndex,]
# test set
heart.test <- heart.data[-trainIndex,]

# check that proportion of positive/negative cases are equal
table(heart.train$HeartDisease) # 0.08559747 prop of positive cases
table(heart.test$HeartDisease) # 0.08558742 prop of positive cases
```



### Data process

```{r}
# X

y= heart.train["HeartDisease"]

# set dimensions of fixed effects

heart.train <- dummy_cols(heart.train, select_columns = 'Smoking')
heart.train <- dummy_cols(heart.train, select_columns = 'AlcoholDrinking')
heart.train <- dummy_cols(heart.train, select_columns = 'Stroke')
heart.train <- dummy_cols(heart.train, select_columns = 'DiffWalking')
heart.train <- dummy_cols(heart.train, select_columns = 'Sex')
heart.train <- dummy_cols(heart.train, select_columns = 'AgeCategory')
heart.train <- dummy_cols(heart.train, select_columns = 'Race')
heart.train <- dummy_cols(heart.train, select_columns = 'Diabetic')
heart.train <- dummy_cols(heart.train, select_columns = 'PhysicalActivity')
heart.train <- dummy_cols(heart.train, select_columns = 'GenHealth')
heart.train <- dummy_cols(heart.train, select_columns = 'Asthma')
heart.train <- dummy_cols(heart.train, select_columns = 'KidneyDisease')
heart.train <- dummy_cols(heart.train, select_columns = 'SkinCancer')


df.train = subset(heart.train, select = -c(Smoking, AlcoholDrinking, Stroke, DiffWalking, Sex, AgeCategory, Race, Diabetic, PhysicalActivity, GenHealth, Asthma, KidneyDisease, SkinCancer))

X = subset(df.train, select = -c(HeartDisease))

```



### logistic regression with elastic net 
```{r}

set.seed(42)
cv_5 = trainControl(method = "cv", number = 5)
fit_elnet = train(
  HeartDisease ~ ., data = df.train,
  method = "glmnet",
  type.measure='class', 
  family='binomial',
  trControl = cv_5,
  tuneLength = 10
)
fit_elnet

```

### Kappa in training dataset 
```{r}
#X = subset(df.train, select = -c(HeartDisease))

yhat <- predict.train(fit_elnet, newdata = X)

confusionMatrix(data= as.factor(yhat), reference=as.factor(df.train$HeartDisease))

```

### ### Kappa in test dataset 
```{r}

heart.test <- dummy_cols(heart.test, select_columns = 'Smoking')
heart.test <- dummy_cols(heart.test, select_columns = 'AlcoholDrinking')
heart.test <- dummy_cols(heart.test, select_columns = 'Stroke')
heart.test <- dummy_cols(heart.test, select_columns = 'DiffWalking')
heart.test <- dummy_cols(heart.test, select_columns = 'Sex')
heart.test <- dummy_cols(heart.test, select_columns = 'AgeCategory')
heart.test <- dummy_cols(heart.test, select_columns = 'Race')
heart.test <- dummy_cols(heart.test, select_columns = 'Diabetic')
heart.test <- dummy_cols(heart.test, select_columns = 'PhysicalActivity')
heart.test <- dummy_cols(heart.test, select_columns = 'GenHealth')
heart.test <- dummy_cols(heart.test, select_columns = 'Asthma')
heart.test <- dummy_cols(heart.test, select_columns = 'KidneyDisease')
heart.test <- dummy_cols(heart.test, select_columns = 'SkinCancer')


df.test = subset(heart.test, select = -c(Smoking, AlcoholDrinking, Stroke, DiffWalking, Sex, AgeCategory, Race, Diabetic, PhysicalActivity, GenHealth, Asthma, KidneyDisease, SkinCancer))

X.test = subset(df.test, select = -c(HeartDisease))

yhatt <- predict.train(fit_elnet, newdata = X.test)

confusionMatrix(data= as.factor(yhatt), reference=as.factor(df.test$HeartDisease))


```

